Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2024-11-27 autovi_paper.qmd
provide more details about the computer vision model

To train a computer vision model a measure of the visible pattern in a plot is needed. We call this the **visual signal strength** (VSS) measuring how prominently a specific set of visual patterns appears in an image. This can be computed for a training set of data, and plots, where the generating distributions are specified.

@@

This distance measure requires knowledge of the true residual distribution, which is typically unknown. Therefore, a computer vision model was used for estimating the residual distribution [see @li2024automated for more details].
XXX this isn't completely clear -- why is the computer vision model used to estimate the residual distribution?  You need to provide at least a bit of the "why" so that this isn't a total non-sequitur. XXX

>>>>

To train a computer vision model, a measure of the visible pattern in a plot is needed. We call this the **visual signal strength** (VSS), which measures how prominently a specific set of visual patterns appears in an image. This can be computed for a training set of data, and plots, where the generating distributions are specified.

@@

This distance measure depends on knowledge of the true residual distribution, which is often unknown. However, it can be estimated by identifying or approximating a mapping from a set of residuals to its corresponding distance measure. The computer vision model presented in @li2024automated approximates this mapping. It is trained on a large number of synthetic regression models, where the data-generating process is known, allowing the distance measure to be explicitly calculated. The model takes a residual plot as input and outputs the corresponding distance measure. Additional details are provided in @li2024automated.

### 2024-11-27 autovi_paper.tex
provide more details about the computer vision model

To train a computer vision model a measure of the visible pattern in a
plot is needed. We call this the \textbf{visual signal strength} (VSS)
measuring how prominently a specific set of visual patterns appears in
an image. This can be computed for a training set of data, and plots,
where the generating distributions are specified.

In the context of regression model diagnostics, it describes the clarity
of visual patterns on a diagnostic plot that may indicate model
violations. This concept can be categorized as weak, moderate, or
strong. However, in our study, we treat it as a continuous positive real
variable. Importantly, its interpretation depends on how it is linked to
a function of the data or the underlying data generating process.
Consequently, the calculation of VSS can be vary across different model
classes or within the same model, depending on the generating function.

VSS is defined as an estimate of a distance measure that quantifies the
disparity between the residual distribution of a fitted classical normal
linear regression model and a reference distribution, and more details
can be found in \citet{li2024automated}. This measure is based on the
Kullback-Leibler (KL) divergence:

@@

This distance measure requires knowledge of the true residual
distribution, which is typically unknown. Therefore, a computer vision
model was used for estimating the residual distribution \citep[see][ for
more details]{li2024automated}.

@@

The main purpose of \texttt{autovi} is to provide rejection decisions
and \(p\)-values for testing the null hypothesis (\(H_0\)) that the
regression model is correctly specified. The package provides automated
interpretation of residual plots using computer vision. The name
\texttt{autovi} stands for \textbf{auto}mated \textbf{v}isual
\textbf{i}nference.

There are two ways to access the package, directly using R or through a
web interface, \texttt{autovi.web}. The web interface has the advantage
that it can be used without installing Python, R and the relevant
packages locally.

\subsection{Motivation for usage}\label{sec-why}

@@

The \texttt{autovi} package is built on the \texttt{bandicoot}
object-oriented programming (OOP) system \citep{bandicoot}, marking a
departure from R's traditional S3 generic system. This OOP architecture
enhances flexibility and modularity, allowing users to redefine key
functions through method overriding. While similar functionality could
be achieved using R's S3 system with generic functions, the OOP
framework offers a more structured and extensible foundation for the
package.

The \texttt{autovi} infrastructure effectively integrates multiple
programming languages and libraries into a comprehensive analytical
tool. It relies on five core libraries from Python and R, each playing a
critical role in the analysis pipeline. In Python, \texttt{pillow}
\citep{clark2015pillow} handles image processing tasks such as reading
and resizing PNG files of residual plots, then converting them into
input tensors for further analysis. The \texttt{TensorFlow}
\citep{abadi2016tensorflow} library, a key component of modern machine
learning, is used to predict the VSS of these plots through a
pre-trained convolutional neural network.

In the R environment, \texttt{autovi} utilizes several libraries.
\texttt{ggplot2} \citep{ggplot2} generates the initial residual plots,
saved as PNG files for visual input. The \texttt{cassowaryr}
\citep{mason2022cassowaryr} library computes scagnostics (scatter plot
diagnostics), providing numerical features that capture statistical
properties of the plots. These scagnostics complement the visual
analysis by offering quantitative metrics as secondary input to the
computer vision model. The \texttt{reticulate} \citep{reticulate}
package bridges R and Python, enabling seamless communication between
the two languages and supporting the integrated infrastructure.

@@

The \texttt{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub.
The code discussed in this paper is based on \texttt{autovi} version
0.4.1.

The package includes internal functions to check the current Python
environment used by the \texttt{reticulate} package. If the necessary
Python packages are not installed in the Python interpreter, an error
will be raised. If you want to select a specific Python environment, you
can do so by calling the \texttt{reticulate::use\_python()} function
before using the \texttt{autovi} package.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Load the \texttt{autovi} package using the \texttt{library()}
  function.
\item
  Create a checker object with a linear regression model.
\item
  Call the \texttt{check()} method of the checker, which, by default,
  predicts the VSS for the true residual plot, 100 null plots, and 100
  bootstrapped plots, storing the predictions internally. A concise
  report of the check results is then printed.
\end{enumerate}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail on the software's
website.

@@

Determining the best way to implement the backend was difficult. In our
initial planning for \texttt{autovi.web}, we considered implementing the
entire web application using the \texttt{webr} framework \citep{webr},
which would have allowed the entire application to run directly in the
user's browser. However, this approach was not feasible at the time of
writing this paper. The reason is that one of the R packages
\texttt{autovi} depends on the R package \texttt{splancs}
\citep{splancs}, which uses compiled Fortran code. A working Emscripten
\citep{zakai2011emscripten} version of this package, which would be
required for \texttt{webr}, was not available.

We also explored the possibility of implementing the web interface using
frameworks built on other languages, such as Python. However, server
hosting domains that natively support Python servers typically do not
have the latest version of R installed. Additionally, calling R from
Python is typically done using the \texttt{rpy2} Python library
\citep{rpy2}, but this approach can be awkward when dealing with
language syntax related to non-standard evaluation. Another option we
considered was renting a server where we could have full control, such
as those provided by cloud platforms like Google Cloud Platform (GCP) or
Amazon Web Services (AWS). However, correctly setting up the server and
ensuring a secure deployment requires significant expertise. Ultimately,
the most practical solution was to use the \texttt{shiny} and
\texttt{shinydashboard} frameworks, which are well-established in the R
community and offer a solid foundation for web application development.

@@

In contrast to \texttt{autovi}, \texttt{autovi.web} does not use the
native Python version of \texttt{TensorFlow}. Instead, it leverages
\texttt{TensorFlow.js}, a JavaScript library that allows the execution
of machine learning models directly in the browser. This choice enables
native browser execution, enhancing compatibility across different user
environments, and shifts the computational load from the server to the
client-side. \texttt{TensorFlow.js} also offers better scalability and
performance, especially when dealing with resource-intensive computer
vision models on the web.

>>>>

To train a computer vision model, a measure of the visible pattern in a
plot is needed. We call this the \textbf{visual signal strength} (VSS),
which measures how prominently a specific set of visual patterns appears
in an image. This can be computed for a training set of data, and plots,
where the generating distributions are specified.

In the context of regression model diagnostics, VSS describes the
clarity of visual patterns on a diagnostic plot that may indicate model
violations. Violations can be categorized as weak, moderate, or strong,
but here we treat it as a continuous positive real variable.
Importantly, its interpretation depends on how it is linked to a
function of the data or the underlying data generating process.
Consequently, the calculation of VSS can be vary across different model
classes or within the same model, depending on the generating function.

VSS is an estimate of the distance between the residual distribution of
a fitted classical normal linear regression model and a reference
distribution; more details can be found in \citet{li2024automated}. The
distance measure is based on the Kullback-Leibler (KL) divergence:

@@

This distance measure depends on knowledge of the true residual
distribution, which is often unknown. However, it can be estimated by
identifying or approximating a mapping from a set of residuals to its
corresponding distance measure. The computer vision model presented in
\citet{li2024automated} approximates this mapping. It is trained on a
large number of synthetic regression models, where the data-generating
process is known, allowing the distance measure to be explicitly
calculated. The model takes a residual plot as input and outputs the
corresponding distance measure. Additional details are provided in
\citet{li2024automated}.

@@

The main purpose of \texttt{autovi} is to provide rejection decisions
and \(p\)-values for testing the null hypothesis (\(H_0\)) that the
regression model is correctly specified. The package provides automated
interpretation of residual plots using computer vision. The name
\texttt{autovi} stands for \textbf{auto}mated \textbf{v}isual
\textbf{i}nference. This functionality can be accessed through the R
package \texttt{autovi}, or through a web interface,
\texttt{autovi.web}, which allows use without the full installation of
R, Python, and package dependencies on the user's system. locally.

\subsection{Motivation}\label{sec-why}

@@

The \texttt{autovi} package is built on the \texttt{bandicoot}
object-oriented programming (OOP) system \citep{bandicoot}, marking a
departure from R's traditional S3 generic system. This OOP architecture
enhances flexibility and modularity, allowing users to redefine key
functions through method overriding XXX I think the technical term is
overloading? XXX. While similar functionality could be achieved using
R's S3 system with generic functions, the OOP framework offers a more
structured and extensible foundation for the package. XXX Not sure if
this last sentence is necessary XXX

The \texttt{autovi} infrastructure effectively integrates multiple
programming languages and libraries into a comprehensive analytical
tool. It relies on five core libraries from Python and R, each playing a
critical role in the analysis pipeline. In Python, \texttt{pillow}
\citep{clark2015pillow} handles image processing tasks such as reading
and resizing PNG files of residual plots, then converting them into
input tensors for further analysis. \texttt{TensorFlow}
\citep{abadi2016tensorflow}, a key component of modern machine learning,
is used to predict the VSS of these plots using a pre-trained
convolutional neural network.

In the R environment, \texttt{autovi} utilizes several libraries.
\texttt{ggplot2} \citep{ggplot2} generates the initial residual plots,
saved as PNG files for visual input. \texttt{cassowaryr}
\citep{mason2022cassowaryr} computes scagnostics (scatter plot
diagnostics), providing numerical features that capture statistical
properties of the plots. These scagnostics complement the visual
analysis by offering quantitative metrics as secondary input to the
computer vision model. \texttt{reticulate} \citep{reticulate} enables
seamless communication between R and Python.

@@

The \texttt{autovi} package is available on CRAN. It is actively
developed and maintained, with the latest updates accessible on GitHub.
This paper uses \texttt{autovi} version 0.4.1. The package includes
internal functions to check the current Python environment used by the
\texttt{reticulate} package. If the necessary Python packages are not
installed in the Python interpreter, an error will be raised. If you
want to select a specific Python environment, you can do so by calling
the \texttt{reticulate::use\_python()} function before using the
\texttt{autovi} package.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Load the \texttt{autovi} package using the \texttt{library()}
  function.
\item
  Create a checker object with a linear regression model.
\item
  Call the \texttt{check()} method of the checker, which, by default,
  predicts the VSS for the true residual plot, 100 null plots, and 100
  bootstrapped plots. The method stores the predictions internally and
  prints a concise results report.
\end{enumerate}

@@

The modules for VSS prediction and \(p\)-value computation are
predefined and cannot be overridden, although users can interact with
them directly through function arguments. Similarly, the image reading
and resizing module is fixed but will adapt to different Keras models by
checking their input shapes. The remaining seven modules are designed to
be overridable, enabling users to tailor the infrastructure to their
specific needs. These modules are discussed in detail in the package
documentation.

@@

Determining the best way to implement the backend was difficult. In our
initial planning for \texttt{autovi.web}, we considered implementing the
entire web application using the \texttt{webr} framework \citep{webr},
which would have allowed the entire application to run directly in the
user's browser. However, \texttt{webr} does not support packages which
use compiled fortran code, which is required by \texttt{splancs}
\citep{splancs}, a dependency of \texttt{autovi}. In the future, it is
possible that a working Emscripten \citep{zakai2011emscripten} version
of this package may allow full \texttt{webr} support.

We also explored the possibility of implementing the web interface using
frameworks built on other languages, such as Python. However, server
hosting domains that natively support Python servers typically do not
have the latest version of R installed. Additionally, calling R from
Python is typically done using the \texttt{rpy2} Python library
\citep{rpy2}, but this approach can be awkward when dealing with
language syntax related to non-standard evaluation. Another option we
considered was renting a server where we could have full control, such
as those provided by cloud platforms like Google Cloud Platform (GCP) or
Amazon Web Services (AWS). However, deploying and maintaining the server
securely requires some expertise. Ultimately, the most practical
solution was to use the \texttt{shiny} and \texttt{shinydashboard}
frameworks, which are well-established in the R community and offer a
solid foundation for web application development.

@@

In contrast to \texttt{autovi}, \texttt{autovi.web} leverages
\texttt{TensorFlow.js}, a JavaScript library that allows the execution
of machine learning models directly in the browser. This choice enables
native browser execution, enhancing compatibility across different user
environments, and shifts the computational load from the server to the
client-side. \texttt{TensorFlow.js} also offers better scalability and
performance, especially when dealing with resource-intensive computer
vision models on the web.
